---
title: "Homework 1"
author: "RuiJun Chen"
date: 2019-09-17
output: github_document
---

# Problem 1

Creating a data frame with random sample of size 8 from a standard Normal distribution, logical vector indicating whether elements of the sample are greater than 0, character vector of length 8, factor vector of length 8, with 3 different factor “levels”
Code chunk with messages off
```{r, message = FALSE}
library(tidyverse)
p1_df = tibble (
  sample = rnorm(8),
  logic_vect = sample > 0,
  char_vect = c("a","b","c","d","e","f","g","h"),
  factor_vect = factor(c("low","medium","high","low","medium","high","low","medium"))
)
```
Attempting to take the mean of each type of vector:

The mean of numeric vector sample is `r mean(pull(p1_df, sample))`

The mean of logical vector logic_vect is `r mean(pull(p1_df, logic_vect))`

The mean of character vector char_vect is `r mean(pull(p1_df, char_vect))`

The mean of factor vector factor_vect is `r mean(pull(p1_df, factor_vect))`

Mean works for numeric and logical vectors but not for charactor or factor vectors, which returns NA instead

Code chunk to convert logical, character, and factor variables to numeric (code chunk shown but no results)
```{r, eval = FALSE}
as.numeric(pull(p1_df, logic_vect))
as.numeric(pull(p1_df, char_vect))
as.numeric(pull(p1_df, factor_vect))
```

Logical vectors convert to 0s and 1s, character vectors throw an error and instead returns NA, and factor vectors convert to numbers for the various 'levels' (in this case 1, 2, 3)

```{r}
pull(p1_df, sample)*as.numeric(pull(p1_df, logic_vect))
pull(p1_df, sample)*as.factor(pull(p1_df, logic_vect))
pull(p1_df, sample)*as.numeric(as.factor((pull(p1_df, logic_vect))))
```


# Problem 2

